```{r,eval=T, echo=F}
klippy::klippy()
```

# Trace and Legend

## Trace 

Trace是圖例legend的最小單位, 圖例的每個項目(legend item)都是一個trace.  
  
沒有圖例表示整個plot是one trace.
```{r}
library(plotly)
# No legend: only one trace
plotly::plot_ly(data=Orange) |>
  add_trace(x=~age, y=~circumference) # no legend: only one trace

# Color creates 5 legend items: 5
plotly::plot_ly(data=Orange) |>
  add_trace(x=~age, y=~circumference, 
    color=~Tree) -> plt_color # 5 legend items: 5 traces
plt_color |> econIDV::showWidget()
```

  * `color` not only serves for color picking purpose, but also serves for traces definition. (Each color is one trace.)
  
  * Legend item ranking is based on:
```{r}
levels(Orange$Tree)
```

### legend interaction

  * Single click (**hide** one): (on existing legend item) remove one trace / (on hidden legend item) restore one trace
  
  * Double click (**show ONLY** one): (on existing legend item) isolate one trace / (on hiddenlegend item) restore all traces
  
**Legend click** serves as a filtering tool.

### color mapping

You can manually define color mapping:
```{r}
# create 5 colors
pal5 = scales::hue_pal()(n=5)
scales::show_col(pal5) 

plotly::plot_ly(data=Orange) |>
  add_trace(x=~age, y=~circumference, 
    color=~Tree, colors=pal5)
```

  * The color mapping is from `levels({color variable})` to `pal5`.

### refine traces

Sometimes we want to split every trace based on some variable. We can use `split`.

When `color` and `split` are both there, traces are created by `color:split` (the interaction of two) whose number of used levels is the number of traces.


```{r}
# split 
plotly::plot_ly(data=Orange) |>
  add_trace(x=~age, y=~circumference, color=~Tree,
    split=~(age>1000)) -> plt_colorSplit # 5 legend items: 5 traces
plt_colorSplit |> econIDV::showWidget()
```


<div class="alert alert-info">
Another way of creating traces is to use `add_xxx` many times.
```{r}
plot_ly(x=~age, y=~circumference) |>
  add_trace(
    data=Orange |> dplyr::filter(Tree==1),
    color=I("blue")
  ) |>
  add_trace(
    data=Orange |> dplyr::filter(Tree==2),
    color=I("green")
  )
```
</div>


### legend item labels

  * determined by trace names.

Create trace names:
```{r}
Orange |> 
  dplyr::mutate(
    trace_name=factor(Tree):factor((age > 1000))
  ) -> Orange
levels(Orange$trace_name)

levels(Orange$trace_name) <- c("3:young", "3:old", "1:young", "1:old", "5:young", "5:old", "2:young", "2:old", "4:young", "4:old")

```

```{r}
plotly::plot_ly(data=Orange) |>
  add_trace(
    x=~age, y=~circumference,
    # color and split define traces (legend item)
    color=~Tree,
    split=~(age>1000),
    # name defines legend item name (label)
    name=~trace_name) -> plt_colorSplitName
plt_colorSplitName |> econIDV::showWidget()
```



## Style

Attributes within an `add_xxx` function can be restyled ad hoc using `plotly::style()`:

```{r}
plotly::plot_ly(data=Orange) |>
  add_trace(x=~age, y=~circumference, 
    color=I("black"), split=~Tree) -> plt0
plt0
```

```{r}
plt0 |>
  plotly::style(
    marker=list(color="red"),
    traces=1:3
  ) 
```

  * trace number is based on legend item order from top to bottom.
  
```{r}
plt0 |> 
  plotly::style(
    marker=list(symbol="diamond")
    ) 
```

  * Without `traces`, it styles all traces.


## Election example

```{r}
examples=econIDV::Examples()
examples$election()
examples$election_env$attach()
```

### data preparation

```{r}
download_electionData()
# prepare choropleth map data
elections |>
  prepare_choroplethMapData() -> df_elections

# 增加 df_elections$得票率顏色說明
df_elections |>
  add_intervalLabels() -> df_elections

df_elections |>
  as.data.frame() |> # to remove tibble class for row.names assignment
  create_sf() |>
  # 移除geometry空的資料, 否則ggplotly會有錯誤
  econIDV::remove_emptyGeometry() ->
  sf_elections
```

### election 2012

```{r sf_election2012}
sf_elections |> dplyr::filter(年=="2012") -> sf_election2012

sf_election2012 |> 
  dplyr::mutate(
    勝黨得票率區間=得票率顏色說明:factor(勝出政黨)
  ) -> sf_election2012

```

### color mapping

```{r , eval=T, echo=F}
teachDS::img_centering2("./img/electionColor.png", "90%") #|> browsable()
```


```{r eval=F, echo=T}
add_trace(
  color= ~..., colors = ---
)
```

  * map `levels(...)` to `---`

```{r}
pal=generate_palette()
sf_election2012 |> 
  plot_ly() |>
  add_sf(
    color=~勝黨得票率區間, 
    colors=pal,
    alpha=1
  ) 
### Something wrong
```


<div class="alert alert-warning">
Notice that we have 10 colors in `pal` even though legend only has 6 colors shown. The length of `pal` must be the same as the length of `levels(勝黨得票率區間)`.
</div>


<div class="alert alert-info">
  * You can use `droplevels()` to drop unused levels if feel necessary.

```{r}
fct = factor(c("b", "b", "c", "a", "a"))
levels(fct)

fct2 = fct[-3] 
levels(fct2)

fct2dropped = droplevels(fct2)
levels(fct2dropped)
```

</div>

```{r}
.intervals = 
  levels(sf_election2012$勝黨得票率區間)

# Check current mapping
econIDV::show_col2(pal, .intervals)
```

We want legend to show from 

  * 0-50:國民黨 to 65-100:國民黨, then
  * 0-50:民進黨 to 65-100:民進黨

Reorder `pal` if necessary:
```{r}
c(rev(pal[1:5]), pal[6:10]) -> twoPartyPals
econIDV::show_col2(twoPartyPals, .intervals)
```


Reorder `.intervals` if necessary:
```{r}
.intervals[c(9,7,5,3,1)]
.intervals[c(
  seq(9,1,-2),
  seq(10,2, -2)
)] -> .intervals2

# Change levels order
sf_election2012$勝黨得票率區間2 <-
  factor(
    sf_election2012$勝黨得票率區間,
    levels=.intervals2
  )

econIDV::show_col2(twoPartyPals,
  levels(sf_election2012$勝黨得票率區間2)
  )
```

```{r}
sf_election2012 |> 
  plot_ly() |>
  add_sf(
    color=~`勝黨得票率區間2`,
    colors=twoPartyPals,
    alpha=1
  ) -> plt_color
plt_color
```

### Legend item labels

Item labels are determined by **trace names**. 

```{r}
sf_election2012$trace_name = 
  sf_election2012$勝黨得票率區間2
levels(sf_election2012$trace_name)
oldLevels =
  levels(sf_election2012$trace_name)

oldLevels |>
  stringr::str_sub(end = -5) -> newLevels
newLevels
newLevels -> 
  levels(sf_election2012$trace_name)
sf_election2012$trace_name
oldLevels
newLevels
```

```{r}
sf_election2012 |> 
  plot_ly() |>
  add_sf(
    name=~trace_name,
    color=~`勝黨得票率區間2`,
    colors=twoPartyPals,
    alpha=1
  ) -> plt_color
plt_color
```

### Legend group

```{r}
sf_election2012 |> 
  plot_ly() |>
  add_sf(
    legendgroup=~勝出政黨,
    name=~trace_name,
    color=~`勝黨得票率區間2`,
    colors=twoPartyPals,
    alpha=1
  ) -> plt_color
plt_color
```

### Legend group titles

```{r , eval=T, echo=F}
teachDS::img_centering2("./img/legendgrouptitle.png", "80%")
```


```{r}
plt_color |>
  plotly::style(
    legendgrouptitle=list(
      text="國民黨"
    ),
    traces=1
  ) |>
  plotly::style(
    legendgrouptitle=list(
      text="民進黨"
    ),
    traces=2
  )
```

<div class="alert alert-warning">

  * `legendgroup` serves as a trace aggregation tool. So we are left with only two traces.

</div>

### Summary


```{r , eval=T, echo=F}
teachDS::img_centering2("./img/legendgroupElection.png", "80%")
```

## Subplot of Election Maps

### Data preparation

```{r}
examples=econIDV::Examples()
examples$election()
examples$election_env$attach()
download_electionData()
# prepare choropleth map data
elections |>
  prepare_choroplethMapData() |>
  # 增加 df_elections$得票率顏色說明
  add_intervalLabels() |>
  dplyr::mutate(
    勝黨得票率區間=得票率顏色說明:factor(勝出政黨)
  ) |>
  # 增加 勝黨得票率區間2
  add_adjustedWinningPartyVoteShare() |>
  # 增加 trace_name
  add_trace_name() -> df_elections

df_elections |>
  as.data.frame() |> # to remove tibble class for row.names assignment
  create_sf() |>
  # 移除geometry空的資料, 否則ggplotly會有錯誤
  econIDV::remove_emptyGeometry() ->
  sf_elections
```

準備勝黨調色盤
```{r}
twoPartyPals <- get_twoPartyPals()
```

標籤格式函數：
```{r}
scales::label_percent(accuracy=0.01, scale=1) -> label_share
```

  * `accuracy`進位到0.01。
  
  * `scale`: 決定原數字要先放大多少倍。如果原數字假設有0.03298代表3.298%, 那`scale=100`。

切出三年資料
```{r}
sf_electionsByYears <- 
  sf_elections |> split(sf_elections$年)
```

### Plotly by year

Each year we plot the graph using:
```{r}
plotly::plot_ly() |>
  add_sf(
    data=sf_electionsByYears[[.x]],
    ...,
    color=~勝黨得票率區間2,
    colors=twoPartyPals,
    alpha=1
  ) 
```

where `...` will be used to explored difference scenarios.

<div class="alert alert-warning">
記得`levels(勝黨得票率區間2)`的長度必需和`twoPartyPals`的長度相同，且其對應關係為所要的對應。
`alpha=1`才能避免內定淡色效果。
</div>

### Group by 地區 


```{r , eval=T, echo=F}
tags$a(
  href="https://tpemartin.github.io/econIDV/election_groupTitleRank.html", target="_blank",teachDS::img_centering2("./img/legendgroupTitleRank.png", "95%"))
```

```{r}
list_splitByCounties <- vector("list", 3)
for(.x in 1:3){
  plotly::plot_ly() |>
    add_sf(
      data=sf_electionsByYears[[.x]],
      legendgroup=~地區,
      name=~label_share(勝出政黨得票率),
      color=~勝黨得票率區間2,
      colors=twoPartyPals,
      alpha=1
    ) -> list_splitByCounties[[.x]]
}

plotly::subplot(list_splitByCounties) -> subplot0
subplot0
```

<div class="alert alert-info">
Legend groups的出現順序是看每一張圖的`name`對應變數的文字排序來決定，由第一張圖開始:
```{r}
sf_electionsByYears[[1]]$地區[
  order(
    label_share(sf_electionsByYears[[1]]$勝出政黨得票率))
] -> .grouptitles
.grouptitles
```

  * 若已包含所有地區則不再看後面圖形的相同規則排序。
</div>

### Adding legendgrouptitles

這裡有19個legendgrouptitle要加上去:

```{r}
subplot0 |> 
  plotly::style(
    legendgrouptitle=list(text=.grouptitles[[1]]),
    traces=1
  ) |> ... |>
  plotly::style(
    legendgrouptitle=list(text=.grouptitles[[19]]),
    traces=19
  )
```

使用迴圈
```{r}
p = subplot0
for(.x in 1:19){
  p=plotly::style(
    p,
    legendgrouptitle=list(text=.grouptitles[[.x]]),
    traces=.x
  )
}
p
```

### Change trace order

If you want trace order to be:
```{r}
c("臺北市", "桃園市", "新竹市", "苗栗縣", "臺中市", 
"彰化縣", "雲林縣", "嘉義市", "臺南市", "高雄市", 
"屏東縣", "基隆市", "新北市", "新竹縣", "宜蘭縣", 
"南投縣", "花蓮縣", "嘉義縣", "臺東縣") -> citiesInOrder
```

You can assign each trace a **legendrank** (0-1000). Smaller number goes to the top.

```{r}
city2rankMapping = 1:19
names(city2rankMapping) <- citiesInOrder
```

```{r}
p |> 
  plotly::style(
    legendrank=city2rankMapping[[.grouptitles[[1]]]],
    traces=1
  ) |> ...|>
  plotly::style(
    legendrank=city2rankMapping[[.grouptitles[[19]]]],
    traces=19
  )   
```

```{r}
for(.x in 1:19){
  p=plotly::style(
    p,
    legendrank=city2rankMapping[[.grouptitles[[.x]]]],
    traces=.x
  )
}
p
```


```{r}
p |>
  plotly::layout(
    legend=list(
      title=list(
        text="得票率"
      ),
      orientation="h"
    )
    )|> econIDV::showWidget()
```

### Complete program

```{r}
list_splitByCounties <- vector("list", 3)
for(.x in 1:3){
  plotly::plot_ly() |>
    add_sf(
      data=sf_electionsByYears[[.x]],
      legendgroup=~地區,
      name=~label_share(勝出政黨得票率),
      color=~勝黨得票率區間2,
      colors=twoPartyPals,
      alpha=1
    ) -> list_splitByCounties[[.x]]
}

plotly::subplot(list_splitByCounties) -> subplot0

city2rankMapping = 1:19
names(city2rankMapping) <- c("臺北市", "桃園市", "新竹市", "苗栗縣", "臺中市", 
"彰化縣", "雲林縣", "嘉義市", "臺南市", "高雄市", 
"屏東縣", "基隆市", "新北市", "新竹縣", "宜蘭縣", 
"南投縣", "花蓮縣", "嘉義縣", "臺東縣")

p = subplot0
for(.x in 1:19){
  p=plotly::style(
    p,
    legendgrouptitle=list(text=.grouptitles[[.x]]),
    legendrank=city2rankMapping[[.grouptitles[[.x]]]],
    traces=.x
  )
}
p |>
  plotly::layout(
    legend=list(
      title=list(
        text="得票率"
      ),
      orientation="h"
    )
    ) -> p
p |> econIDV::showWidget()
```


```{r,echo=FALSE}
htmltools::save_html(p, "/Users/martinl/Github/econIDV/docs/election_groupTitleRank.html")
```

### Group by 勝出政黨 

```{r}
plotly::plot_ly() |>
    add_sf(
      data=sf_electionsByYears[[1]],
      legendgroup=~勝出政黨,
      name=~得票率顏色說明,
      # split=~地區,
      color=~勝黨得票率區間2, colors=twoPartyPals,
      alpha=1
    ) |>
  plotly::style(
    legendgrouptitle=list(
      text="國民黨"
    ),
    traces=1
  ) |>
  plotly::style(
    legendgrouptitle=list(
      text="民進黨"
    ),
    traces=2
  )
```


```{r}
list_splitByCountiesGroupByParties <- vector("list", 3)
for(.x in 1:3){
  plotly::plot_ly() |>
    add_sf(
      data=sf_electionsByYears[[.x]],
      legendgroup=~勝出政黨,
      name=~label_share(勝出政黨得票率),
      split=~地區,
      color=~勝黨得票率區間2, colors=twoPartyPals,
      alpha=1
    ) -> list_splitByCountiesGroupByParties[[.x]]
}

plotly::subplot(list_splitByCountiesGroupByParties) -> 
  subplot_splitByCountiesGroupByParties
subplot_splitByCountiesGroupByParties |>
  plotly::style(
    legendgrouptitle=list(text=htmlEscape("國民黨")),
    traces=1
  ) |>
  plotly::style(
    legendgrouptitle=list(text="民進黨"),
    traces=2
  ) |>
  plotly::layout(
    legend=list(
      title=list(
        text="得票率"
      ),
      orientation="h"
    )
    ) -> subplot_splitByCountiesGroupByParties
subplot_splitByCountiesGroupByParties |> econIDV::showWidget()
```

## Hide elements

### via css

```{r}
tagList(
  `widget`,
  tags$style("css styling rules")
)
```

```{r}
tagList(
  subplot_splitByCountiesGroupByParties,
  tags$style(
    "g.traces:not(:first-child),rect.scrollbar {display: none;}"
  )
)
```

```{r}
tagList(
  subplot_splitByCountiesGroupByParties,
  htmltools::includeCSS("data/widget.css")
) |> econIDV::showWidget()
```

```{r}
subplot_splitByCountiesGroupByParties |>
  tagList(
    tags$style(
    "g.traces:not(:first-child),rect.scrollbar {display: none;}"
  )) |> econIDV::showWidget()
```


### legendgroup by 地區

```{r}
sf_electionsByYears |>
  plot_allYears(
    legendgroup=~地區,
    split=~paste(地區,年, sep = ":")) -> list_plts
subplot(list_plts) -> subplot0
```

```{r}
subplot0 |>
  econIDV::get_ranked_splitNames() -> ranked_splits
subplot0 |>
  econIDV::style_groupTitles_basedOnRankedSplitNames(from=1, to=3) -> subplot1

subplot1 |>
  plotly::layout(
    legend=list(font=list(size=1, color="#00000000"), orientation="h")
  ) |> econIDV::showWidget()
```


```{r echo=F}
sf_electionsByYears |>
  plot_allYears(
    legendgroup=~勝出政黨,
    split=~paste(地區,年, sep = ":")) -> list_plts1
subplot(list_plts1) -> subplot1
subplot1 |> econIDV::showWidget()
```

### and sharedData by 地區

SharedData links row name across plots. A row name is used to identify **ONE** trace. 

Legend (ungrouped one) item each represents a trace.

Since both click on plot and on legend are dealing with filtering out **ONE** trace at a time. They should be consistent.

  * What used in 


```{r}
examples=econIDV::Examples()
examples$election()
examples$election_env$attach()
download_electionData()
# prepare choropleth map data
elections |>
  prepare_choroplethMapData() -> df_elections

# create new column for split
#  * can represents each color category
#  * its levels sequence is adequate. 
df_elections |>
  create_newLevels4split() -> df_elections$色標

levels(df_elections$色標)
levels(df_elections$取色得票率區間)

df_elections |>
  as.data.frame() |> # to remove tibble class for row.names assignment
  create_sf() |>
  # 移除geometry空的資料, 否則ggplotly會有錯誤
  econIDV::remove_emptyGeometry() ->
  sf_elections
```


```{r}
sf_electionsByYears <- 
  sf_elections |> split(sf_elections$年)
```


```{r}
# 設定consistent row names, 再轉成sharedData  
sharedSf = vector("list", 3)
for(.x in 1:3){
  row.names(sf_electionsByYears[[.x]]) <- 
  sf_electionsByYears[[.x]]$地區
  sharedSf[[.x]] <- 
    crosstalk::SharedData$new(
      sf_electionsByYears[[.x]], group="elections"
    )
}
```

row name 為地區

```{r}
list_plts <- vector("list",3)
for(.x in 1:3){
  plotly::plot_ly() |>
    add_sf(
      data=sharedSf[[.x]],
      legendgroup=~地區,
      split=~paste(地區,年, sep = ":"),
      color=~I(as.character(得票率顏色)),
      alpha=1
    ) -> list_plts[[.x]]
}

subplot(list_plts) -> subplot0
subplot0 |> econIDV::showWidget()
```

```{r}
subplot0 |>
  econIDV::get_ranked_splitNames() -> ranked_splits
subplot0 |>
  econIDV::style_groupTitles_basedOnRankedSplitNames(from=1, to=3) -> subplot1
```


```{r}
sf_electionsByYears[[1]] |> View()
```


```{r}
subplot1 |>
  plotly::highlight(
    off="plotly_doubleclick"
  ) |>
  plotly::style(
    name="50%",
    traces=1
  ) |>
  plotly::layout(
    legend=list(
      # font=list(size=1, color="#00000000"),
      orientation="v",
      itemclick=F) 
  ) |> econIDV::showWidget()
```

  * `itemclick=F` prohibit click on item event (double click on legend group event still works).
  
